\section{Data and design}\label{sec:data}

\paragraph{Cohorts.} Every cohort is reduced to a common per-chain schema
(one CDR3 with its V/J and epitope) and a paired schema
($\mathrm{CDR3}_\alpha,V_\alpha,J_\alpha,\mathrm{CDR3}_\beta,V_\beta,J_\beta$, epitope).
The cohorts span the expected signal-to-noise (S/N) range:

\begin{center}\small
\begin{tabular}{@{}lll@{}}
\toprule
cohort & definition & role \\
\midrule
TCRvdb true  & TCRvdb, $p_{\text{adj}}<10^{-5}$ & experimentally verified (top) \\
VDJdb HQ     & VDJdb human MHC-I, $\ge 2$ references per (TCR, epitope) & strongly supported \\
VDJdb LQ     & VDJdb human MHC-I, single reference & weakly supported \\
TCRvdb false & TCRvdb, $p_{\text{adj}}\ge 10^{-5}$ & experimentally refuted \\
immrep25 pos & immrep25 \texttt{label}=1 & \textbf{system under test} \\
AIRR random & pooled real repertoire (isalgo/airr\_control), random labels & real-data floor \\
AIRR non-random & 20 largest SRA donors; epitope $=$ a donor's top-50 clonotypes & structured real control \\
OLGA random  & raw OLGA generation, random labels & noise floor \\
MLR expanded ($\beta$) & real proliferating TCR$\beta$ clones; epitope $=$ a replicate's top-200 & MIRA-analogue control \\
\bottomrule
\end{tabular}
\end{center}

\noindent Cohorts draw on VDJdb~\cite{Shugay_2017}, the functionally-validated
TCRvdb/MATCHMAKERS database~\cite{Messemaker_2025}, the immrep25 unseen-peptide
benchmark~\cite{Richardson2026}, and an OLGA control built from the TCR generation
model of Murugan~et~al.~\cite{Murugan_2012}. The immrep25 test set covers 20
nine-mer peptides that are \emph{absent from VDJdb}, with 50 positives (and 450
negatives) each; only positives are audited.
VDJdb HQ is defined per chain (TRA and TRB separately) as a (TCR, epitope) supported
by $\ge 2$ distinct references, with a paired complex counted HQ if \emph{either}
chain qualifies. TCRvdb true/false are dominated by the same two epitopes
(YLQPRTFLL, GLCTLVAML), giving an epitope-matched verified-vs-refuted contrast.

\paragraph{One-vs-many.} Every metric is computed \emph{per epitope} against all
others (each epitope with $\ge 30$ records vs.\ the rest of its cohort) and then
aggregated to a dataset value---geometric mean $\pm$ 95\% CI for homology S/N, mean
$\pm$ 95\% CI for pairing---across epitopes. The signal therefore attaches to
individual epitopes and the reported error bars are the epitope-to-epitope spread
($n$ = number of qualifying epitopes); no bootstrap resampling is used.

\paragraph{A structured real-data control.} Random synthetic labels give S/N${=}1$ by
construction, so we add a control whose epitope grouping is \emph{real and structured}:
the \textbf{AIRR non-random} cohort. From the isalgo/airr\_benchmark SRA collection we
take the 20 largest samples that carry $\ge 50$ TRA and $\ge 50$ TRB clonotypes; each
\emph{donor} contributes its top 50 TRA and top 50 TRB clonotypes (by read count) as one
virtual epitope. This mimics how a real epitope-specific dataset is built (expanded
clones per sample), but the ``epitopes'' are individual donors, not antigens. Two
random-label floors (OLGA random and the pooled AIRR random) are kept for reference. The
20 donors are listed in \texttt{results/airr\_donors.tsv}.

\paragraph{A control that mimics immrep25's own methodology.} immrep25's positive
$\beta$ labels come from a MIRA assay that identifies \emph{expanded} (proliferating)
TCR$\beta$ clones responding to a peptide pool~\cite{Richardson2026}; the $\alpha$ chain
is added afterward by separate pairing. We therefore build a $\beta$-only analogue,
\textbf{MLR expanded}, from real mixed-lymphocyte-reaction proliferating repertoires
(isalgo/airr\_benchmark \texttt{alice/mlr})---genuinely antigen-driven expanded clones.
The six proliferating samples are in fact three independent reactions, each captured in
two heavily overlapping duplicate samples (sharing $\sim$50--65\% of their top clones),
with two sequencing replicates apiece. We take each replicate's top-200 TCR$\beta$
clonotypes (by read count) as one virtual epitope---four per reaction, twelve in
all---and, because same-reaction replicates share the same expanded clones, score each
one-vs-many against the \emph{other reactions} only (never its own siblings). This is the
closest possible real-data mimic of how immrep25's positives were generated.

\paragraph{Pipeline.}
\begin{center}
\begin{tikzpicture}[node distance=6mm and 10mm,>=Latex,
  box/.style={draw,rounded corners,align=center,font=\footnotesize,inner sep=3pt,minimum height=8mm},
  data/.style={box,fill=blue!6}, proc/.style={box,fill=orange!8}, result/.style={box,fill=green!8}]
\node[data] (imm) {immrep25\\TCRvdb\\VDJdb};
\node[proc,right=of imm] (coh) {normalize\\cohorts};
\node[proc,above right=3mm and 9mm of coh] (airr) {AIRR non-random\\(20 donors)};
\node[proc,below right=3mm and 9mm of coh] (olga) {random floors\\(AIRR, OLGA)};
\node[proc,right=32mm of coh] (hom) {homology\\S/N};
\node[proc,below=8mm of hom] (pair) {pairing\\$z$};
\node[result,right=10mm of hom] (fig) {S/N\\hierarchy};
\draw[->] (imm)--(coh);
\draw[->] (coh)--(airr); \draw[->] (coh)--(olga);
\draw[->] (airr)--(hom); \draw[->] (olga)--(pair);
\draw[->] (coh)--(hom); \draw[->] (coh)--(pair);
\draw[->] (hom)--(fig); \draw[->] (pair.east)-|(fig);
\end{tikzpicture}
\end{center}
